---
title: "Data Sources and Overview"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

```{r setup, include = FALSE}
library(tidyverse)
library(readr)
library(plotly)

knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.align = 'center')

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Data Sources

The source of the dataset comes from R for Data Science's [Tidy Tuesday Project](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-03-09), which provides weekly releases of raw datasets for users to wrangle and analyze. The data itself originates from [FiveThirtyEight](https://github.com/fivethirtyeight/data/tree/master/bechdel), containing movies ranging from 1970 to 2013, and merges data from several sources:

* **[IMDB](https://imdb.com)**: Provides information about each film, such as year, runtime, genres, accolades, fan and critic ratings. 
* **[BechdelTest.com](http://bechdeltest.com)**: The site is user-operated database where fans can rate films based on the criteria of the Bechdel Test. The site has detailed, coded information about the year, title, and whether it passes each level of the Bechdel Test requirements (see [Data Cleaning](Data_Sources.html#Data_cleaning) for further information about scoring).
* **[The-Numbers.com](https://The-Numbers.com)**: Contains financial information for films, including the budget and gross domestic and international revenues. 

## Data Import and Cleaning

The dataset was distributed as a CSV file, `movies.csv`. Let's load and inspect the variables in the raw data.

```{r load}
movies = read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/movies.csv')

head(movies) %>% 
  knitr::kable(digits = 3) %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), font_size = 12) %>% 
  kableExtra::scroll_box(width = "100%", height = "300px")
```
  
<br>

The `movies.csv` dataset contains `r nrow(movies)` observations and `r ncol(movies)` variables. Each observation corresponds to a film, and contains variables for title, IMDB ID, year, genres, ratings, budget, domestic and international gross revenue, and 2013-adjusted budget and revenues. The variables `clean_test` categorizes movies according to the different Bechdel Test criteria whether it passes the Bechdel Test (based on qualitative scores and a binary pass/fail),

### Data cleaning

The data cleaning steps involve the following:

- Converting character to numeric variables for `domgross`, `intgross`, `domgross_2013`, `intgross_2013`, and `runtime`
- Recoding and relevelling the numeric `decade_code` variable into a `decade` factor variable according to the decade of release
- Creating an `award_winner` variable that is `TRUE` if the movie has won at least 1 Oscar, Golden Globe, or BAFTA award
- Splitting the `genre` character variable into dummy-coded variables indicating whether the movie falls under one of 20 different genres
- Creating a `bechdel_score` categorical variable of 4 levels corresponding to the Bechdel Test criteria:
  - Less than 2 women
  - Women don't talk to each other
  - Women only talk about men
  - Dubious (Passes Bechdel Test, but is disputed)
  - Passes Bechdel Test
- Creating a logical variable `pass_bechdel` that indicates whether the movie passes or fails the Bechdel Test based on the `bechdel_score` variable, which is:
  - `TRUE` if the movie is coded as "Dubious" or "Passes Bechdel Test"
  - `FALSE` if the movie is coded as "Less than 2 women", "Women don't talk to each other", or "Women only talk about men"
- Creating variables for profit and ROI for each movie, adjusted to 2013 inflation. For this, we will use worldwide box office revenue instead of domestic revenue.
  - `profit` = `intgross_2013 – budget_2013`
  - `ROI` = `profit / budget_2013`
- Renaming key variables to relevant names
- Removing extraneous variables

The table below displays the full tidied data, and summarizes each variable within the dataset.

```{r tidy movies}
movies_df = movies %>% 
  mutate(test = as.factor(test), 
         clean_test = fct_recode(clean_test, 
                                 "Less than 2 women" = "nowomen", 
                                 "Don't talk to each other" = "notalk",
                                 "Only talk about men" = "men", 
                                 "Dubious" = "dubious",
                                 "Passes Bechdel" = "ok"),
         clean_test = fct_relevel(clean_test, c("Less than 2 women", "Don't talk to each other", 
                                                "Only talk about men", "Dubious", "Passes Bechdel")),
         binary = ifelse(clean_test == "Dubious" | clean_test == "Passes Bechdel", TRUE, FALSE), 
         domgross = as.numeric(domgross),
         intgross = as.numeric(intgross),
         domgross_2013 = as.numeric(domgross_2013),
         intgross_2013 = as.numeric(intgross_2013),
         decade_code = case_when(year >= 1970 & year < 1980 ~ "1970-1979",
                            year >= 1980 & year < 1990 ~ "1980-1989",
                            year >= 1990 & year < 2000 ~ "1990-1999",
                            year >= 2000 & year < 2010 ~ "2000-2009",
                            year >= 2010 & year < 2020 ~ "2010-2013"),
         decade_code = as.factor(decade_code),
         title = str_replace(title, "&#39;", "'"),
         title = str_replace(title, "&amp;", "&"),
         title = str_replace(title, "&agrave;", "à"),
         title = str_replace(title, "&aring;", "å"),
         title = str_replace(title, "&auml;", "ä"), 
         runtime = as.numeric(str_replace(runtime, " min", "")),
         award_winner = ifelse(str_detect(awards, "Won") & (str_detect(awards, "Golden Globe") | 
                                                              str_detect(awards, "Oscar") | str_detect(awards, "BAFTA")), T, F), 
         profit = intgross_2013 - budget_2013,
         ROI = profit/budget_2013) %>% 
  separate(genre, into = c("g1", "g2", "g3"), sep = ", ") %>% 
  rename("pass_bechdel" = binary,
         "bechdel_score" = clean_test,
         "decade" = decade_code) %>% 
  mutate(
    action = ifelse(g1 == "Action" | g2 == "Action" | g3 == "Action", TRUE, FALSE),
    adventure = ifelse(g1 == "Adventure" | g2 == "Adventure" | g3 == "Adventure", TRUE, FALSE),
    animation = ifelse(g1 == "Animation" | g2 == "Animation" | g3 == "Animation", TRUE, FALSE),
    biography = ifelse(g1 == "Biography" | g2 == "Biography" | g3 == "Biography", TRUE, FALSE),
    comedy = ifelse(g1 == "Comedy" | g2 == "Comedy" | g3 == "Comedy", TRUE, FALSE),
    crime = ifelse(g1 == "Crime" | g2 == "Crime" | g3 == "Crime", TRUE, FALSE),
    documentary = ifelse(g1 == "Documentary" | g2 == "Documentary" | g3 == "Documentary", TRUE, FALSE),
    drama = ifelse(g1 == "Drama" | g2 == "Drama" | g3 == "Drama", TRUE, FALSE),
    family = ifelse(g1 == "Family" | g2 == "Family" | g3 == "Family", TRUE, FALSE),
    fantasy = ifelse(g1 == "Fantasy" | g2 == "Fantasy" | g3 == "Fantasy", TRUE, FALSE),
    history = ifelse(g1 == "History" | g2 == "History" | g3 == "History", TRUE, FALSE),
    horror = ifelse(g1 == "Horror" | g2 == "Horror" | g3 == "Horror", TRUE, FALSE),
    music = ifelse(g1 == "Music" | g2 == "Music" | g3 == "Music", TRUE, FALSE),
    musical = ifelse(g1 == "Musical" | g2 == "Musical" | g3 == "Musical", TRUE, FALSE),
    mystery = ifelse(g1 == "Mystery" | g2 == "Mystery" | g3 == "Mystery", TRUE, FALSE),
    romance = ifelse(g1 == "Romance" | g2 == "Romance" | g3 == "Romance", TRUE, FALSE),
    sci_fi = ifelse(g1 == "Sci-Fi" | g2 == "Sci-Fi" | g3 == "Sci-Fi", TRUE, FALSE),
    sport = ifelse(g1 == "Sport" | g2 == "Sport" | g3 == "Sport", TRUE, FALSE),
    thriller = ifelse(g1 == "Thriller" | g2 == "Thriller" | g3 == "Thriller", TRUE, FALSE),
    war = ifelse(g1 == "War" | g2 == "War" | g3 == "War", TRUE, FALSE),
    western = ifelse(g1 == "Western" | g2 == "Western" | g3 == "Western", TRUE, FALSE)
  ) %>% 
  mutate(across(action:western, ~replace_na(., FALSE))) %>%
  select(year, title, bechdel_score, pass_bechdel, budget_2013:intgross_2013, decade, imdb_id, metascore, imdb_rating, award_winner, runtime, profit, ROI, action:western) 

readr::write_csv(movies_df, "./movies_df.csv")

movies_df %>% 
  knitr::kable(digits = 3) %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), font_size = 12) %>% 
  kableExtra::scroll_box(width = "100%", height = "300px")
```

```{r summarise df}
summary(movies_df) %>% 
  knitr::kable(digits = 3) %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), font_size = 12) %>% 
  kableExtra::scroll_box(width = "100%", height = "300px")
```

  
<br>

Our resulting dataset contains `r nrow(movies_df)` observations and `r ncol(movies_df)` variables, with information about the movie's decade, genre, and runtime, a Bechdel Test categorical variable, a binary Bechdel Test variable, and information about its budget, revenue, and ratings. This cleaned and aggregated data will serve the basis of our exploratory and formal analyses.

## Superhero Bonus

The `Superheroes` dataset was obtained using the Selector Gadget Chrome Extension. 
<br>
We created a custom [IMDB](https://www.imdb.com/list/ls566366989/) list containing all released movies from Marvel Cinematic Universe, as well as one for the DC Extended Universe. Several DC Superheroes movies not in the Extended Universe were also added. Using the Selector Gadget, we pulled variables `title` and `year`. 
<br>
Data on budget, domestic and international gross was pulled using the Selector Gadget from [Wikipedia](https://en.wikipedia.org/wiki/List_of_Marvel_Cinematic_Universe_films).
<br>
Data on the Bechdel test results for each movie was merged from our `movies_df` dataframe, as well as [bechdeltest.com](https://bechdeltest.com/). 
<br></br>
The datasets obtained from each website were cleaned and merged to create the Superheroes.xlsx file used for the Superheroes dashboard. 



