

---
title: "Exploratory Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
date: "2022-12-05"
---

```{r setup, include = FALSE}
library(tidyverse)
library(readr)
library(plotly)
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.align = 'center')



theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r load, echo=FALSE}
raw_bechdel = read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/raw_bechdel.csv')
movies = read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/movies.csv')
```



```{r tidy movies, echo =FALSE}
movies_df = movies %>% 
  mutate(test = as.factor(test), 
         clean_test = fct_recode(clean_test, 
                                 "Less than 2 women" = "nowomen", 
                                 "Don't talk to each other" = "notalk",
                                 "Only talk about men" = "men", 
                                 "Dubious" = "dubious",
                                 "Passes Bechdel" = "ok"),
         clean_test = fct_relevel(clean_test, c("Less than 2 women", "Don't talk to each other", 
                                                "Only talk about men", "Dubious", "Passes Bechdel")),
         binary = ifelse(clean_test == "Dubious" | clean_test == "Passes Bechdel", TRUE, FALSE), 
         domgross = as.numeric(domgross),
         intgross = as.numeric(intgross),
         domgross_2013 = as.numeric(domgross_2013),
         intgross_2013 = as.numeric(intgross_2013),
         decade_code = case_when(year >= 1970 & year < 1980 ~ "1970-1979",
                            year >= 1980 & year < 1990 ~ "1980-1989",
                            year >= 1990 & year < 2000 ~ "1990-1999",
                            year >= 2000 & year < 2010 ~ "2000-2009",
                            year >= 2010 & year < 2020 ~ "2010 - present"),
         decade_code = as.factor(decade_code),
         title = str_replace(title, "&#39;", "'"),
         title = str_replace(title, "&amp;", "&"),
         title = str_replace(title, "&agrave;", "à"),
         title = str_replace(title, "&aring;", "å"),
         title = str_replace(title, "&auml;", "ä"), 
         runtime = as.numeric(str_replace(runtime, " min", "")),
         award_winner = ifelse(str_detect(awards, "Won") & (str_detect(awards, "Golden Globe") | 
                                                              str_detect(awards, "Oscar") | str_detect(awards, "BAFTA")), T, F), 
         profit = intgross_2013 - budget_2013,
         ROI = profit/budget_2013) %>% 
  separate(genre, into = c("g1", "g2", "g3"), sep = ", ") %>% 
  rename("pass_bechdel" = binary,
         "bechdel_score" = clean_test,
         "decade" = decade_code) %>% 
  mutate(
    action = ifelse(g1 == "Action" | g1 == "Action" | g1 == "Action", TRUE, FALSE),
    adventure = ifelse(g1 == "Adventure" | g1 == "Adventure" | g1 == "Adventure", TRUE, FALSE),
    animation = ifelse(g1 == "Animation" | g1 == "Animation" | g1 == "Animation", TRUE, FALSE),
    biography = ifelse(g1 == "Biography" | g1 == "Biography" | g1 == "Biography", TRUE, FALSE),
    comedy = ifelse(g1 == "Comedy" | g1 == "Comedy" | g1 == "Comedy", TRUE, FALSE),
    crime = ifelse(g1 == "Crime" | g1 == "Crime" | g1 == "Crime", TRUE, FALSE),
    documentary = ifelse(g1 == "Documentary" | g1 == "Documentary" | g1 == "Documentary", TRUE, FALSE),
    drama = ifelse(g1 == "Drama" | g1 == "Drama" | g1 == "Drama", TRUE, FALSE),
    family = ifelse(g1 == "Family" | g1 == "Family" | g1 == "Family", TRUE, FALSE),
    fantasy = ifelse(g1 == "Fantasy" | g1 == "Fantasy" | g1 == "Fantasy", TRUE, FALSE),
    history = ifelse(g1 == "History" | g1 == "History" | g1 == "History", TRUE, FALSE),
    horror = ifelse(g1 == "Horror" | g1 == "Horror" | g1 == "Horror", TRUE, FALSE),
    music = ifelse(g1 == "Music" | g1 == "Music" | g1 == "Music", TRUE, FALSE),
    musical = ifelse(g1 == "Musical" | g1 == "Musical" | g1 == "Musical", TRUE, FALSE),
    mystery = ifelse(g1 == "Mystery" | g1 == "Mystery" | g1 == "Mystery", TRUE, FALSE),
    romance = ifelse(g1 == "Romance" | g1 == "Romance" | g1 == "Romance", TRUE, FALSE),
    sci_fi = ifelse(g1 == "Sci-Fi" | g1 == "Sci-Fi" | g1 == "Sci-Fi", TRUE, FALSE),
    sport = ifelse(g1 == "Sport" | g1 == "Sport" | g1 == "Sport", TRUE, FALSE),
    thriller = ifelse(g1 == "Thriller" | g1 == "Thriller" | g1 == "Thriller", TRUE, FALSE),
    war = ifelse(g1 == "War" | g1 == "War" | g1 == "War", TRUE, FALSE),
    western = ifelse(g1 == "Western" | g1 == "Western" | g1 == "Western", TRUE, FALSE)
  ) %>% 
  select(year, title, bechdel_score, pass_bechdel, budget_2013:intgross_2013, decade, imdb_id, language, metascore, imdb_rating, award_winner, runtime, profit, ROI, action:western) 

readr::write_csv(movies_df, "./movies_df.csv")


```



## linear models

We tested the association of ` categorical Bechdel score` with outcomes `budget`, `profit`, `IMDB rating` and `Meta scores` individually using linear regression.  

Bechdel score of `"less than 2 women"` was taken as the reference category.
A movie with < than 2 women would fail all 3 dimensions of the Bechdel test which are 1) have at least 2 women 2) women talk to each other 3) they need to talk about something other than a man. Thus the intercept will be used to compare movies that fail all dimensions of the Bechdel test with the movies that pass all dimensions of the  the Bechdel test.

### Linear regression of budget vs. categorical Bechdel score



```{r}
budget_dist_p = movies_df %>% ggplot(aes(x = budget_2013)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "Budget ($, 2013-adjusted)",
    title = "Distribution of movie budgets")

ggplotly(budget_dist_p)
```

Since budget was heavily right skewed, we log-transformed the data instead to enforce more normal distributions. We quantitatively tested the association of log of budget with categorical Bechdel score by regressing budget against the categories. 

```{r budget}

log_budget= movies_df %>%
  lm(log(budget_2013)~bechdel_score, data=.) %>% 
  broom::tidy()

log_budget


```
Based on parameter estimate of -0.31003 with a p value of 0.01, 
Passing all 3 Bechdel dimensions had a significant effect and a negative relationship with the Budget of the movies.

Movies that failed every Bechdel dimension had an average budget of $33,963,942
Movies that passed all 3 Bechdel dimensions had an average budget of $24,910,750

### Linear regression of profit vs. categorical Bechdel score

```{r}
profit_dist = movies_df %>% ggplot(aes(x = profit)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "Profit",
    title = "Distribution of Profits")

ggplotly(profit_dist)
```


In the graph, profit appears to be heavily right skewed. Thus we log-transformed the data to enforce more normal distributions. Additionally, before the logarithm is applied, 1 was added to the base value to prevent applying a logarithm to a 0 value. This will be subtracted while calculating the results.

We quantitatively tested the association of profit with categorical Bechdel score by regressing profit against the categories adjusting for budget


```{r}

profit = movies_df %>% 
  lm(log(profit+1)~bechdel_score+log(budget_2013), data =.)

summary(profit)

```
Based on the parameter estimate of -0.044, 
Bechdel scores don't significantly influence Profit and passing all 3 dimensions of the Bechdel test had a negative relationship with Profit. However since the p-value is 0.7 > 0.05, the difference was found to be statistically insignificant


### Linear regression of IMDB rating vs. categorical Bechdel score

…
```{r}
imdb = movies_df %>% ggplot(aes(x = imdb_rating)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "IMDB rating",
    title = "Distribution of IMDB")

ggplotly(imdb)
```
The graph shows an approximately normal distribution.

Testing association of IMDB rating with categorical Bechdel scores using linear regression

```{r}
imdb = movies_df %>% 
  lm(imdb_rating~bechdel_score, data=.)

summary(imdb)
```

Based on the Parameter estimate of -0.344 with a p value of 0.00019,
movies that pass all the dimensions of Bechdel test received 0.34 point lower IMDB rating than movies that fail all the dimensions of a Bechdel test.

### Linear regression of Meta score vs. categorical Bechdel score

```{r}
meta = movies_df %>% ggplot(aes(x = metascore)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "meta score",
    title = "Distribution of Metascore")

ggplotly(meta)
```
The graph shows an approximately normal distribution.

Testing association of meta score with categorical Bechdel scores using linear regression

```{r}
meta = movies_df %>% 
  lm(metascore~bechdel_score, data=.) 

summary(meta)

```
Based on the parameter estimate -0.4642,  movies that pass all dimensions of the Bechdel test had a negative relationship with Meta scores. However since the p-value 0.7 > 0.05, the difference was found to be statistically insignificant

Movies that pass all the dimensions of Bechdel test received 0.46 points less in meta scores than movies that fail all the dimensions of a Bechdel test.

