

---
title: "Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
date: "2022-12-05"
---

```{r setup, include = FALSE}
library(tidyverse)
library(readr)
library(plotly)
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.align = 'center')



theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r load, echo=FALSE}
raw_bechdel = read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/raw_bechdel.csv')
movies = read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/movies.csv')
```




```{r tidy movies, echo=FALSE}
movies_df = movies %>% 
  mutate(test = as.factor(test), 
         clean_test = fct_recode(clean_test, 
                                 "Less than 2 women" = "nowomen", 
                                 "Don't talk to each other" = "notalk",
                                 "Only talk about men" = "men", 
                                 "Dubious" = "dubious",
                                 "Passes Bechdel" = "ok"),
         clean_test = fct_relevel(clean_test, c("Less than 2 women", "Don't talk to each other", 
                                                "Only talk about men", "Dubious", "Passes Bechdel")),
         binary = ifelse(clean_test == "Dubious" | clean_test == "Passes Bechdel", TRUE, FALSE), 
         domgross = as.numeric(domgross),
         intgross = as.numeric(intgross),
         domgross_2013 = as.numeric(domgross_2013),
         intgross_2013 = as.numeric(intgross_2013),
         decade_code = case_when(year >= 1970 & year < 1980 ~ "1970-1979",
                            year >= 1980 & year < 1990 ~ "1980-1989",
                            year >= 1990 & year < 2000 ~ "1990-1999",
                            year >= 2000 & year < 2010 ~ "2000-2009",
                            year >= 2010 & year < 2020 ~ "2010-2013"),
         decade_code = as.factor(decade_code),
         title = str_replace(title, "&#39;", "'"),
         title = str_replace(title, "&amp;", "&"),
         title = str_replace(title, "&agrave;", "à"),
         title = str_replace(title, "&aring;", "å"),
         title = str_replace(title, "&auml;", "ä"), 
         runtime = as.numeric(str_replace(runtime, " min", "")),
         award_winner = ifelse(str_detect(awards, "Won") & (str_detect(awards, "Golden Globe") | 
                                                              str_detect(awards, "Oscar") | str_detect(awards, "BAFTA")), T, F), 
         profit = intgross_2013 - budget_2013,
         ROI = profit/budget_2013) %>% 
  separate(genre, into = c("g1", "g2", "g3"), sep = ", ") %>% 
  rename("pass_bechdel" = binary,
         "bechdel_score" = clean_test,
         "decade" = decade_code) %>% 
  mutate(
    action = ifelse(g1 == "Action" | g2 == "Action" | g3 == "Action", TRUE, FALSE),
    adventure = ifelse(g1 == "Adventure" | g2 == "Adventure" | g3 == "Adventure", TRUE, FALSE),
    animation = ifelse(g1 == "Animation" | g2 == "Animation" | g3 == "Animation", TRUE, FALSE),
    biography = ifelse(g1 == "Biography" | g2 == "Biography" | g3 == "Biography", TRUE, FALSE),
    comedy = ifelse(g1 == "Comedy" | g2 == "Comedy" | g3 == "Comedy", TRUE, FALSE),
    crime = ifelse(g1 == "Crime" | g2 == "Crime" | g3 == "Crime", TRUE, FALSE),
    documentary = ifelse(g1 == "Documentary" | g2 == "Documentary" | g3 == "Documentary", TRUE, FALSE),
    drama = ifelse(g1 == "Drama" | g2 == "Drama" | g3 == "Drama", TRUE, FALSE),
    family = ifelse(g1 == "Family" | g2 == "Family" | g3 == "Family", TRUE, FALSE),
    fantasy = ifelse(g1 == "Fantasy" | g2 == "Fantasy" | g3 == "Fantasy", TRUE, FALSE),
    history = ifelse(g1 == "History" | g2 == "History" | g3 == "History", TRUE, FALSE),
    horror = ifelse(g1 == "Horror" | g2 == "Horror" | g3 == "Horror", TRUE, FALSE),
    music = ifelse(g1 == "Music" | g2 == "Music" | g3 == "Music", TRUE, FALSE),
    musical = ifelse(g1 == "Musical" | g2 == "Musical" | g3 == "Musical", TRUE, FALSE),
    mystery = ifelse(g1 == "Mystery" | g2 == "Mystery" | g3 == "Mystery", TRUE, FALSE),
    romance = ifelse(g1 == "Romance" | g2 == "Romance" | g3 == "Romance", TRUE, FALSE),
    sci_fi = ifelse(g1 == "Sci-Fi" | g2 == "Sci-Fi" | g3 == "Sci-Fi", TRUE, FALSE),
    sport = ifelse(g1 == "Sport" | g2 == "Sport" | g3 == "Sport", TRUE, FALSE),
    thriller = ifelse(g1 == "Thriller" | g2 == "Thriller" | g3 == "Thriller", TRUE, FALSE),
    war = ifelse(g1 == "War" | g2 == "War" | g3 == "War", TRUE, FALSE),
    western = ifelse(g1 == "Western" | g2 == "Western" | g3 == "Western", TRUE, FALSE)
  ) %>% 
  mutate(across(action:western, ~replace_na(., FALSE))) %>%
  select(year, title, bechdel_score, pass_bechdel, budget_2013:intgross_2013, decade, imdb_id, metascore, imdb_rating, award_winner, runtime, profit, ROI, action:western) 

readr::write_csv(movies_df, "./movies_df.csv")

```




## linear models

We tested the association of ` categorical Bechdel score` with outcomes `budget`, `profit`, `IMDB rating` and `Meta scores` individually using linear regression.  

Bechdel score of `"less than 2 women"` was taken as the reference category.
A movie with < than 2 women would fail all 3 dimensions of the Bechdel test which are 1) have at least 2 women 2) women talk to each other 3) they need to talk about something other than a man. 
Thus the intercept will be used to compare movies that fail all dimensions of the Bechdel test with the movies that pass all dimensions of the Bechdel test.

### Linear regression of budget vs. categorical Bechdel score



```{r}
budget = movies_df %>% ggplot(aes(x = budget_2013)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "Budget ($, 2013-adjusted)",
    title = "Distribution of movie budgets")

ggplotly(budget)

```

Since budget was heavily `right skewed`, we log-transformed it to enforce more normal distributions.

```{r}
log_budget = movies_df %>% ggplot(aes(x = log(budget_2013))) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "Budget ($, 2013-adjusted)",
    title = "Distribution of movie budgets")

ggplotly(log_budget)
```

We quantitatively tested the association of log of budget with categorical Bechdel score using linear regression 

```{r budget}

log_budget= movies_df %>%
  lm(log(budget_2013)~bechdel_score, data=.)%>% 
  broom::tidy()
  
knitr::kable(log_budget, digits=3)

```

Based on parameter estimate of -0.31 with a p value of 0.01, 
passing all 3 Bechdel dimensions had a negative relationship with the Budget of the movies. Movies that passed all 3 Bechdel dimensions had a lower budget than movies that failed every Bechdel dimension.

* Movies that failed every Bechdel dimension had an average budget of `$33,963,942`
* Movies that passed all 3 Bechdel dimensions had an average budget of `$24,910,750`

### Linear regression of profit vs. categorical Bechdel score

```{r}
profit = movies_df %>% ggplot(aes(x = profit)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "Profit",
    title = "Distribution of Profits")

ggplotly(profit)
```


In the graph, profit appears to be heavily `right skewed.` Thus we log-transformed the data to enforce more normal distributions. Additionally, before the logarithm is applied, 1 was added to the base value to prevent applying a logarithm to a 0 value. This will be subtracted while calculating the results.

```{r}
profit = movies_df %>% ggplot(aes(x = log(profit))) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "Profit",
    title = "Distribution of Profits")

ggplotly(profit)
```
We quantitatively tested the association of profit with categorical Bechdel score by regressing profit against the categories adjusting for budget


```{r}

profit = movies_df %>% 
  lm(log(profit+1)~bechdel_score, data =.) %>% 
  broom::tidy()


knitr::kable(profit, digits=3)

profit

```
Based on the parameter estimate of -0.044, 
Profits of movies that passed all 3 dimensions of the Bechdel test had lower profits than movies that failed all dimensions of the Bechdel test. However since the p-value is 0.7 > 0.05, the difference was not found to be statistically significant

* Movies that failed all dimensions of the Bechdel test made an average profit of `$75,526,942`
* Movies that passed all dimensions of the Bechdel test made an average profit of `$57,082,034`


### Linear regression of IMDB rating vs. categorical Bechdel score

…
```{r}
imdb = movies_df %>% ggplot(aes(x = imdb_rating)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "IMDB rating",
    title = "Distribution of IMDB")

ggplotly(imdb)
```
* The graph shows an approximately `normal distribution.`

Testing association of IMDB rating with categorical Bechdel scores using linear regression

```{r}
imdb = movies_df %>% 
  lm(imdb_rating~bechdel_score, data=.) %>% 
  broom::tidy()

knitr::kable(imdb, digits=3)

```

Based on the Parameter estimate of -0.344 with a p value of 0.00019, movies that passed all 3 dimesions of Bechdel test had a negative relationship with IMDB ratings.

* Movies that passed all dimensions of the Bechdel test received `0.34 points` lower IMDB ratings than movies that failed all dimensions of the Bechdel test.

### Linear regression of Metascore vs. categorical Bechdel score

```{r}
meta = movies_df %>% ggplot(aes(x = metascore)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "meta score",
    title = "Distribution of Metascore")

ggplotly(meta)
```
* The graph shows an approximately `normal distribution.`

Testing association of Metascore with categorical Bechdel scores using linear regression

```{r}
meta = movies_df %>% 
  lm(metascore~bechdel_score, data=.) %>% 
  broom::tidy()
knitr::kable(meta, digits=3)

```
Based on the parameter estimate -0.4642,  movies that pass all dimensions of the Bechdel test had a negative relationship with Meta scores. However since the p-value 0.7 > 0.05, the difference was not statistically significant

* Movies that passed all dimensions of the Bechdel test received `0.46 points` less in Metascore than movies that failed all dimensions of the Bechdel test.



