---
title: "Regression Modeling"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

```{r setup, include=FALSE}
library(tidyverse)
library(MASS)
library(readr)
library(plotly)

knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.align = 'center')

theme_set(theme_minimal() + theme(legend.position = "bottom"))

# Read in dataset
movies_df <- read_csv("../Cathy/movies_df.csv")
```

## Linear Models

We tested the association of `bechdel_score` with outcomes `budget`, `profit`, `imdb_rating`, and `metascore` individually using linear regression.  

Bechdel score of `"Less than 2 women"` was taken as the reference category.
A movie with less than 2 women would fail all three dimensions of the Bechdel test which are: 
  1. have at least 2 women in the movie
  2. the women talk to each other 
  3. they talk about something other than a man. 
Thus, the intercept will be used to compare movies that fail all dimensions of the Bechdel test with the movies that pass all dimensions of the Bechdel test.

### Linear Regression of Budget vs. Categorical Bechdel Score

```{r}
budget = movies_df %>% ggplot(aes(x = budget_2013)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Budget (Dollars, 2013-adjusted)" ,
    y = "Count",
    title = "Distribution of Movie Budgets")

ggplotly(budget)

```

Since budget was heavily right-skewed, we log-transformed it to enforce more normal distributions.

```{r}
log_budget = movies_df %>% ggplot(aes(x = log(budget_2013))) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x =  "Log-Budget (Dollars, 2013-adjusted)",
    y = "Count" ,
    title = "Distribution of Log-Transformed Movie Budgets")

ggplotly(log_budget)
```

We then quantitatively tested the association of the `log(budget)` variable with the categorical Bechdel score using linear regression.

Model statement:
$log(Budget) =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r budget}

log_budget= movies_df %>%
  lm(log(budget_2013)~bechdel_score, data=.)%>% 
  broom::tidy()

  
knitr::kable(log_budget, digits=3)

```

Based on parameter estimate of -0.31 with a p value of 0.01, 
passing all three Bechdel dimensions had a negative relationship with the budget of the movies. Movies that passed all three Bechdel dimensions had a lower budget than movies that failed every Bechdel dimension.

* Movies that failed every Bechdel dimension had an average budget of `$33,963,942`
* Movies that passed all three Bechdel dimensions had an average budget of `$24,910,750`

### Linear Regression of Profit vs. Categorical Bechdel Score

```{r}
profit = movies_df %>% ggplot(aes(x = profit)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Profit (Dollars)",
    y = "Count",
    title = "Distribution of Profits")

ggplotly(profit)
```


In the graph, profit appears to be heavily right skewed. Thus, we log-transformed the data to enforce more normal distributions. Additionally, before the logarithm is applied, 1 was added to the base value to prevent applying a logarithm to a 0 value. This will be subtracted while calculating the results.

```{r}
profit = movies_df %>% ggplot(aes(x = log(profit))) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Log-Profit",
    y = "Count",
    title = "Distribution of Log-Transformed Profits")

ggplotly(profit)
```
We quantitatively tested the association of `profit` with the categorical Bechdel score using linear regression.

 
Model statement : 
$log(profit) =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r}

profit = movies_df %>% 
  lm(log(profit+1)~bechdel_score, data =.) %>% 
  broom::tidy()


knitr::kable(profit, digits=3)
```
Based on the parameter estimate of -0.044, 
profits of movies that passed all three dimensions of the Bechdel test had lower profits than movies that failed all dimensions of the Bechdel test. However since the p-value of 0.7 > 0.05, the difference was not found to be statistically significant.

* Movies that failed all dimensions of the Bechdel test made an average profit of `$75,526,942`
* Movies that passed all dimensions of the Bechdel test made an average profit of `$57,082,034`


### Linear Regression of IMDB Rating vs. Categorical Bechdel Score

…
```{r}
imdb = movies_df %>% ggplot(aes(x = imdb_rating)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "IMDB Rating",
    y = "Count",
    title = "Distribution of IMDB Ratings")

ggplotly(imdb)
```
* The graph shows an approximately normal distribution.

Testing association of IMDB rating with categorical Bechdel scores using linear regression:

Model statement:
$IMDB =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r}
imdb = movies_df %>% 
  lm(imdb_rating~bechdel_score, data=.) %>% 
  broom::tidy()


knitr::kable(imdb, digits=3)

```

Based on the parameter estimate of -0.344 with a p-value of 0.00019, movies that passed all three dimensions of the Bechdel test had a negative relationship with `imdb_rating`.

* Movies that passed all dimensions of the Bechdel test received 0.34 points lower IMDB ratings than movies that failed all dimensions of the Bechdel test.

### Linear Regression of Metascore vs. Categorical Bechdel Score

```{r}
meta = movies_df %>% ggplot(aes(x = metascore)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Metascore",
    y = "Count",
    title = "Distribution of Metascores")

ggplotly(meta)
```
* The graph shows an approximately normal distribution.

Testing association of Metascore with categorical Bechdel scores using linear regression:

Model statement: 
$Metascore =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r}
meta = movies_df %>% 
  lm(metascore~bechdel_score, data=.) %>% 
  broom::tidy()
knitr::kable(meta, digits=3)

```
Based on the parameter estimate -0.4642, movies that pass all dimensions of the Bechdel test had a negative relationship with Metascores. However since the p-value of 0.7 > 0.05, the difference was not statistically significant.

* Movies that passed all dimensions of the Bechdel test received 0.46 points less in Metascore than movies that failed all dimensions of the Bechdel test.



