---
title: "Exploratory Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
date: "2022-12-05"
---

```{r setup, include = FALSE}
library(tidyverse)
library(readr)
library(plotly)

knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.align = 'center')

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Data Sources and Overview 

The source of the dataset comes from R for Data Science's [Tidy Tuesday Project](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-03-09), which provides weekly releases of raw datasets for users to wrangle and analyze. The data itself originates from FiveThirtyEight, containing movies ranging from 1970 to 2013, and merges data from several sources:

* **[IMDB](https://imdb.com)**: Provides information about each film, such as year, runtime, genre, language, rating.
* **[BechdelTest.com](http://bechdeltest.com)**: Provides year, title, and a Bechdel Test score ranging from 0-3.
* **[The-Numbers.com](https://The-Numbers.com)**: Provides information about budget and revenue of each movie.

## Data Import and Cleaning

```{r load}
raw_bechdel = read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/raw_bechdel.csv')

movies = read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/movies.csv')
```

The dataset was distributed in two files, `movies.csv`, `raw_bechdel.csv`.

* `movies.csv`: A data file of `r nrow(movies)` observations and `r ncol(movies)` variables, which includes movies from 1970 to 2013, and contains variables for title, IMDB ID, year, Bechdel test (uncleaned, cleaned, binary pass/fail), genres, ratings, budget, domestic and international gross revenue, and 2013-adjusted budget and revenues.
* `raw_bechdel.csv`: A data file of `r nrow(raw_bechdel)` observations and `r ncol(raw_bechdel)` variables, containing movies from 1888 to 2021, and contains variables for title, movie ID, IMDB ID, and raw Bechdel test score (0 to 3).

### Data cleaning

The data cleaning steps involve the following:

- Converting character to numeric variables for `domgross`, `intgross`, `domgross_2013`, `intgross_2013`, and `runtime`
- Converting the `binary` character variable to a logical variable `pass_bechdel` that indicates whether the movie passes or fails the Bechdel Test
- Recoding and relevelling the numeric `decade_code` variable into a `decade` factor variable according to the decade of release
- Creating an `award_winner` variable that is `TRUE` if the movie has won at least 1 Oscar, Golden Globe, or BAFTA award
- Splitting the `genre` character variable into dummy-coded variables indicating whether the movie falls under one of 20 different genres
- Renaming key variables to relevant names
- Recoding and relevelling the `clean_test` variable into 4 levels corresponding to the Bechdel Test criteria:
  - 0 = Less than 2 women
  - 1 = Women don't talk to each other
  - 2 = Women only talk about men
  - 3 = Dubious (uncertain whether it passes Bechdel Test)
  - 4 = Passes Bechdel Test
- Creating variables for profit and ROI for each movie, adjusted to 2013 inflation. For this, we will use worldwide box office revenue instead of domestic revenue.
  - `profit` = `intgross_2013 – budget_2013`
  - `ROI` = `profit / budget_2013`

```{r tidy movies}
movies_df = movies %>% 
  mutate(test = as.factor(test), 
         clean_test = fct_recode(clean_test, 
                                 "Less than 2 women" = "nowomen", 
                                 "Don't talk to each other" = "notalk",
                                 "Only talk about men" = "men", 
                                 "Dubious" = "dubious",
                                 "Passes Bechdel" = "ok"),
         clean_test = fct_relevel(clean_test, c("Less than 2 women", "Don't talk to each other", 
                                                "Only talk about men", "Dubious", "Passes Bechdel")),
         binary = ifelse(binary == "PASS", TRUE, FALSE), 
         domgross = as.numeric(domgross),
         intgross = as.numeric(intgross),
         domgross_2013 = as.numeric(domgross_2013),
         intgross_2013 = as.numeric(intgross_2013),
         decade_code = case_when(year >= 1970 & year < 1980 ~ "1970-1979",
                            year >= 1980 & year < 1990 ~ "1980-1989",
                            year >= 1990 & year < 2000 ~ "1990-1999",
                            year >= 2000 & year < 2010 ~ "2000-2009",
                            year >= 2010 & year < 2020 ~ "2010 - present"),
         decade_code = as.factor(decade_code),
         title = str_replace(title, "&#39;", "'"),
         title = str_replace(title, "&amp;", "&"),
         title = str_replace(title, "&agrave;", "à"),
         title = str_replace(title, "&aring;", "å"),
         title = str_replace(title, "&auml;", "ä"), 
         runtime = as.numeric(str_replace(runtime, " min", "")),
         award_winner = ifelse(str_detect(awards, "Won") & (str_detect(awards, "Golden Globe") | 
                                                              str_detect(awards, "Oscar") | str_detect(awards, "BAFTA")), T, F), 
         profit = intgross_2013 - budget_2013,
         ROI = profit/budget_2013) %>% 
  separate(genre, into = c("g1", "g2", "g3"), sep = ", ") %>% 
  rename("pass_bechdel" = binary,
         "bechdel_score" = clean_test,
         "decade" = decade_code) %>% 
  mutate(
    action = ifelse(g1 == "Action" | g1 == "Action" | g1 == "Action", TRUE, FALSE),
    adventure = ifelse(g1 == "Adventure" | g1 == "Adventure" | g1 == "Adventure", TRUE, FALSE),
    animation = ifelse(g1 == "Animation" | g1 == "Animation" | g1 == "Animation", TRUE, FALSE),
    biography = ifelse(g1 == "Biography" | g1 == "Biography" | g1 == "Biography", TRUE, FALSE),
    comedy = ifelse(g1 == "Comedy" | g1 == "Comedy" | g1 == "Comedy", TRUE, FALSE),
    crime = ifelse(g1 == "Crime" | g1 == "Crime" | g1 == "Crime", TRUE, FALSE),
    documentary = ifelse(g1 == "Documentary" | g1 == "Documentary" | g1 == "Documentary", TRUE, FALSE),
    drama = ifelse(g1 == "Drama" | g1 == "Drama" | g1 == "Drama", TRUE, FALSE),
    family = ifelse(g1 == "Family" | g1 == "Family" | g1 == "Family", TRUE, FALSE),
    fantasy = ifelse(g1 == "Fantasy" | g1 == "Fantasy" | g1 == "Fantasy", TRUE, FALSE),
    history = ifelse(g1 == "History" | g1 == "History" | g1 == "History", TRUE, FALSE),
    horror = ifelse(g1 == "Horror" | g1 == "Horror" | g1 == "Horror", TRUE, FALSE),
    music = ifelse(g1 == "Music" | g1 == "Music" | g1 == "Music", TRUE, FALSE),
    musical = ifelse(g1 == "Musical" | g1 == "Musical" | g1 == "Musical", TRUE, FALSE),
    mystery = ifelse(g1 == "Mystery" | g1 == "Mystery" | g1 == "Mystery", TRUE, FALSE),
    romance = ifelse(g1 == "Romance" | g1 == "Romance" | g1 == "Romance", TRUE, FALSE),
    sci_fi = ifelse(g1 == "Sci-Fi" | g1 == "Sci-Fi" | g1 == "Sci-Fi", TRUE, FALSE),
    sport = ifelse(g1 == "Sport" | g1 == "Sport" | g1 == "Sport", TRUE, FALSE),
    thriller = ifelse(g1 == "Thriller" | g1 == "Thriller" | g1 == "Thriller", TRUE, FALSE),
    war = ifelse(g1 == "War" | g1 == "War" | g1 == "War", TRUE, FALSE),
    western = ifelse(g1 == "Western" | g1 == "Western" | g1 == "Western", TRUE, FALSE)
  ) %>% 
  select(year, title, bechdel_score, pass_bechdel, budget_2013:intgross_2013, decade, imdb_id, language, metascore, imdb_rating, award_winner, runtime, profit, ROI, action:western) 

readr::write_csv(movies_df, "./movies_df.csv")

head(movies_df) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("striped", "hover", full_width = F) %>% 
  kableExtra::scroll_box(width = "100%", height = "300px")
```

Our resulting dataset contains `r nrow(movies_df)` observations and `r ncol(movies_df)` variables, indicating information basic information about the movie's decade, genre, and runtime, whether it passes the Bechdel Test, and information about its budget, revenue, and ratings.

## Exploratory Data Analysis

### Distribution of Bechdel Test Scores

```{r}
movies_df %>% 
  group_by(bechdel_score) %>% 
  summarise(N = n()) %>% 
  mutate(Proportion = N/sum(N)) %>% 
  rename("Bechdel Test Criterion" = bechdel_score) %>% 
  knitr::kable(digits = 3) %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

### Bechdel Test Scores by decade

```{r}
less2_df = movies_df %>% 
  filter(bechdel_score == "Less than 2 women") %>% 
  group_by(decade, bechdel_score) %>% 
  summarise(n = n()) 

notalk_df = movies_df %>% 
  filter(bechdel_score == "Don't talk to each other") %>% 
  group_by(decade, bechdel_score) %>% 
  summarise(n = n()) 

talkmen_df = movies_df %>% 
  filter(bechdel_score == "Only talk about men") %>% 
  group_by(decade, bechdel_score) %>% 
  summarise(n = n()) 

dubious_df = movies_df %>% 
  filter(bechdel_score == "Dubious") %>% 
  group_by(decade, bechdel_score) %>% 
  summarise(n = n()) 

pass_df = movies_df %>% 
  filter(bechdel_score == "Passes Bechdel") %>% 
  group_by(decade, bechdel_score) %>% 
  summarise(n = n()) 

table = bind_cols(less2_df, notalk_df, talkmen_df, dubious_df, pass_df)

table %>% 
  plot_ly(x = ~decade...1, y = ~n...3, type = "bar", name = "Less than 2 women",
          marker = list(color = "darkred")) %>% 
  add_trace(y = ~n...6, name = "Don't talk to each other", marker = list(color = "red")) %>% 
  add_trace(y = ~n...9, name = "Only talk about men", marker = list(color = "darkorange")) %>% 
  add_trace(y = ~n...12, name = "Dubious", marker = list(color = "yellow")) %>% 
  add_trace(y = ~n...15, name = "Passes Bechdel", marker = list(color = "lightgreen")) %>% 
  layout(barmode = "stack",
         xaxis = list(title = "Decade"),
         yaxis = list(title = "Count"))
```

### Bechdel Test Scores by genre


### Bechdel Test by IMDB Rating

```{r}
movies_df %>% 
  plot_ly(y = ~imdb_rating, x = ~bechdel_score, type = "scatter", 
          mode = "markers", marker = list(color = ~imdb_rating)) %>% 
  layout(yaxis = list(title = list(text = "IMDB Rating", standoff = 5), tickfont = list(size = 10), gridcolor = "white"),
         xaxis = list(title = "Bechdel Criterion"), tickfont = list(size = 10), gridcolor = "gray")
```



### Distribution of budgets according to Bechdel scores

Next, we want to explore how movie budgets may differ according to the primacy of women's roles in movies. Grouping by Bechdel score, we can compute the median budget, adjusted to 2013 inflation. 


```{r}
movies_df %>% 
  group_by(bechdel_score) %>% 
  summarise(median_budget = median(budget_2013)) %>% 
  plot_ly(x = ~median_budget, y = ~bechdel_score, type = "bar", color = ~bechdel_score, colors = "YlGn") %>% 
  layout(yaxis = list(title = "Bechdel Criterion", standoff = 10),
         xaxis = list(title = "Median movie budget ($)"), 
         legend = list(reverse = TRUE))

movies_df %>% 
  plot_ly(x = ~bechdel_score, y = ~budget_2013, type = "box", text = ~title) %>% 
  layout(yaxis = list(title = "Movie budget ($)", standoff = 10),
         xaxis = list(title = "Bechdel Criterion" ), 
         legend = list(reverse = TRUE))
```

We can  visualize the median budgets with a bar chart to see that movies featuring two women who don't talk to each other appear to have much larger budgets than the rest. Movies that pass the Bechdel test also appear to have slightly smaller budgets than movies that don't pass.

Next steps...

- F-test (ANOVA) of differences in budget


```{r}
budget_dist_p = movies_df %>% ggplot(aes(x = budget_2013)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Count",
    y = "Budget ($, 2013-adjusted)",
    title = "Distribution of movie budgets")

ggplotly(budget_dist_p)
```

It appears that budget is heavily right-skewed. For this, we will need to run a Kruskal-Wallace test in place of an F-test. Below are the results:

```{r}
kruskal.test(budget_2013 ~ bechdel_score, data = movies_df) %>% 
  broom::tidy() %>% 
  rename("Test statistic" = statistic,
         "p-value" = p.value, 
         "Parameter (df)" = parameter,
         "Method" = method) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::kable_styling(font_size = 12)

```

```{r}

```

