---
title: "Regression Modeling"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

```{r setup, include=FALSE}
library(tidyverse)
library(MASS)
library(readr)
library(plotly)

knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.align = 'center')

theme_set(theme_minimal() + theme(legend.position = "bottom"))

# Read in dataset
movies_df <- read_csv("./Cathy/movies_df.csv")
```

## Linear Models

We tested the association of `bechdel_score` with outcomes `budget`, `profit`, `imdb_rating`, and `metascore` individually using linear regression.  

Bechdel score of `"Less than 2 women"` was taken as the reference category.
A movie with less than 2 women would fail all three dimensions of the Bechdel test which are: 
  1. have at least 2 women in the movie
  2. the women talk to each other 
  3. they talk about something other than a man. 
Thus, the intercept will be used to compare movies that fail all dimensions of the Bechdel test with the movies that pass all dimensions of the Bechdel test.

### Linear Regression of Budget vs. Categorical Bechdel Score

```{r}
budget = movies_df %>% ggplot(aes(x = budget_2013)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Budget (Dollars, 2013-adjusted)" ,
    y = "Count",
    title = "Distribution of Movie Budgets")

ggplotly(budget)

```

Since budget was heavily right-skewed, we log-transformed it to enforce more normal distributions.

```{r}
log_budget = movies_df %>% ggplot(aes(x = log(budget_2013))) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x =  "Log-Budget (Dollars, 2013-adjusted)",
    y = "Count" ,
    title = "Distribution of Log-Transformed Movie Budgets")

ggplotly(log_budget)
```

We then quantitatively tested the association of the `log(budget)` variable with the categorical Bechdel score using linear regression.

Model statement:
$log(Budget) =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r budget}

log_budget= movies_df %>%
  lm(log(budget_2013)~bechdel_score, data=.)%>% 
  broom::tidy()

  
knitr::kable(log_budget, digits=3)

```

Based on parameter estimate of -0.31 with a p value of 0.01, 
passing all three Bechdel dimensions had a negative relationship with the budget of the movies. Movies that passed all three Bechdel dimensions had a lower budget than movies that failed every Bechdel dimension.

* Movies that failed every Bechdel dimension had an average budget of `$33,963,942`
* Movies that passed all three Bechdel dimensions had an average budget of `$24,910,750`

### Linear Regression of Profit vs. Categorical Bechdel Score

```{r}
profit = movies_df %>% ggplot(aes(x = profit)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Profit (Dollars)",
    y = "Count",
    title = "Distribution of Profits")

ggplotly(profit)
```


In the graph, profit appears to be heavily right skewed. Thus, we log-transformed the data to enforce more normal distributions. Additionally, before the logarithm is applied, 1 was added to the base value to prevent applying a logarithm to a 0 value. This will be subtracted while calculating the results.

```{r}
profit = movies_df %>% ggplot(aes(x = log(profit))) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Log-Profit",
    y = "Count",
    title = "Distribution of Log-Transformed Profits")

ggplotly(profit)
```
We quantitatively tested the association of `profit` with the categorical Bechdel score using linear regression.

 
Model statement : 
$log(profit) =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r}

profit = movies_df %>% 
  lm(log(profit+1)~bechdel_score, data =.) %>% 
  broom::tidy()


knitr::kable(profit, digits=3)
```
Based on the parameter estimate of -0.044, 
profits of movies that passed all three dimensions of the Bechdel test had lower profits than movies that failed all dimensions of the Bechdel test. However since the p-value of 0.7 > 0.05, the difference was not found to be statistically significant.

* Movies that failed all dimensions of the Bechdel test made an average profit of `$75,526,942`
* Movies that passed all dimensions of the Bechdel test made an average profit of `$57,082,034`


### Linear Regression of IMDB Rating vs. Categorical Bechdel Score

…
```{r}
imdb = movies_df %>% ggplot(aes(x = imdb_rating)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "IMDB Rating",
    y = "Count",
    title = "Distribution of IMDB Ratings")

ggplotly(imdb)
```
* The graph shows an approximately normal distribution.

Testing association of IMDB rating with categorical Bechdel scores using linear regression:

Model statement:
$IMDB =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r}
imdb = movies_df %>% 
  lm(imdb_rating~bechdel_score, data=.) %>% 
  broom::tidy()


knitr::kable(imdb, digits=3)

```

Based on the parameter estimate of -0.344 with a p-value of 0.00019, movies that passed all three dimensions of the Bechdel test had a negative relationship with `imdb_rating`.

* Movies that passed all dimensions of the Bechdel test received 0.34 points lower IMDB ratings than movies that failed all dimensions of the Bechdel test.

### Linear Regression of Metascore vs. Categorical Bechdel Score

```{r}
meta = movies_df %>% ggplot(aes(x = metascore)) + geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Metascore",
    y = "Count",
    title = "Distribution of Metascores")

ggplotly(meta)
```
* The graph shows an approximately normal distribution.

Testing association of Metascore with categorical Bechdel scores using linear regression:

Model statement: 
$Metascore =\beta_0+ \beta_1(<2 women) +\beta_2(don't talk) +\beta_3(talk about men) + beta_4(dubious)+beta_5(pass)$

```{r}
meta = movies_df %>% 
  lm(metascore~bechdel_score, data=.) %>% 
  broom::tidy()
knitr::kable(meta, digits=3)

```
Based on the parameter estimate -0.4642, movies that pass all dimensions of the Bechdel test had a negative relationship with Metascores. However since the p-value of 0.7 > 0.05, the difference was not statistically significant.

* Movies that passed all dimensions of the Bechdel test received 0.46 points less in Metascore than movies that failed all dimensions of the Bechdel test.


## Stepwise Regression Modeling 

In order to assess whether a movie passing the Bechdel test is a significant predictor of that movie's success, we utilized a stepwise selection algorithm. The `stepAIC` command from the `MASS` package performs stepwise model selection by optimizing AIC.

The setup of the regression models here follows the methodology outlined previously including log-transforming variables like `profit` and `budget_2013` to enforce a normal distribution. Additionally, the binary Bechdel test variable was applied in these models to optimize degrees of freedom.

We assessed movie success based on three factors: `profit`, `imdb_rating`, and `budget_2013`.

### Profit

We first assessed the Bechdel test as a potential predictor of `profit`. Other potential predictors included in the model were `budget_2013`, `imdb_rating`, and all genre variables. 

```{r profit log transfrom, warning=FALSE}
# Stepwise regression for profit
## logProfit ~ Bechdel (binary) + logBudget + IMDB + Genre
modelprofit = lm(log(profit +1) ~ pass_bechdel + log(budget_2013) + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
stepprofit <- stepAIC(modelprofit, direction = "both", trace = FALSE) %>% broom::tidy()
knitr::kable(stepprofit, digits = 3)
```
The algorithm selected budget, IMDB rating, and the genre variables of adventure, biography, drama, horror, music, romance, sci-fi, and western as influential predictors of a movie's profit. `pass_bechdel` was not selected as an influential predictor.

### IMDB Ratings

The second metric we assessed as a measure of movie success was movie reviews as measured by `imdb_rating`. Other potential predictors we included in the model were budget, `runtime`, `profit`, `award_winner`, and all genre variables.

```{r imdb, warning=FALSE}
# Stepwise regression for imdb ratings
## IMDB ~ Bechdel (binary) + logBudget + runtime + award_winner + profit + Genre
modelIMDB = lm(imdb_rating ~ pass_bechdel + log(budget_2013) + profit + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
stepIMDB <- stepAIC(modelIMDB, direction = "both", trace = FALSE) %>% broom::tidy()
knitr::kable(stepIMDB, digits = 3)
```
The algorithm selected the Bechdel test, budget, profit, and the genre variables of action, adventure, animation, biography, comedy, crime, drama, family, fantasy, horror, music, romance, thriller, and war as influential variables of a movie's IMDB rating. The p-value of `pass_bechdel` is significant at 5% and, on average, movies that pass the Bechdel test have a 0.255 point lower rating on IMDB than movies that do not pass, when adjusting for budget, profit, and the aforementioned genre variables.

### Budget

As budget is a significant predictor of both profit and IMDB ratings, we also assessed whether passing the Bechdel test is an influential predictor of a movie's budget. Other potential predictors we included in the model were `runtime` and all genre variables.

```{r budget log transform}
# Stepwise regression for budget
## logBudget ~ Bechdel (binary) + runtime + Genre
modelbudget = lm(log(budget_2013) ~ pass_bechdel + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
stepbudget <- stepAIC(modelbudget, direction = "both", trace = FALSE) %>%  broom::tidy()
knitr::kable(stepbudget, digits = 3)
```
The algorithm selected `runtime` and the genre variables of action, adventure, animation, biography, comedy, crime, documentary, drama, family, fantasy, musical, mystery, romance, sci-fi, sport, and thriller. `pass_bechdel` was not selected as an influential predictor.


In conclusion, based on our data, whether or not a movie passes the Bechdel test is not a significant predictor of that movie's potential for success based on the criteria of profit and budget. It is a significant predictor of a movie's IMDB ratings where movies that pass have, on average, a lower rating than movies that do not pass. Therefore, passing the Bechdel test can be considered detrimental to a movie's success.

