---
title: "Stepwise Regression"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

```{r setup, include=FALSE}
library(tidyverse)
library(MASS)
knitr::opts_chunk$set(echo = TRUE)
# Read in dataset
movies_df <- read_csv("../Cathy/movies_df.csv")
```

## Using Stepwise Selection for Linear Regression: Is the Bechdel test a significant predictor of movie success?

In order to assess whether a movie passing the Bechdel test is a significant predictor of that movie's success, we utilized a stepwise selection algorithm. The `stepAIC` command from the `MASS` package performs stepwise model selection by optimizing AIC.

The setup of the regression models here follows the methodology outlined previously including log-transforming variables like `profit` and `budget_2013` to enforce a normal distribution. Additionally, the binary Bechdel test variable was applied in these models to optimize degrees of freedom.

We assessed movie success based on three factors: `profit`, `imdb_rating`, and `budget_2013`.

### Profit

We first assessed the Bechdel test as a potential predictor of `profit`. Other potential predictors included in the model were `budget_2013`, `imdb_rating`, and all genre variables. 

```{r profit log transfrom, warning=FALSE}
# Stepwise regression for profit
## logProfit ~ Bechdel (binary) + logBudget + IMDB + Genre
modelprofit = lm(log(profit +1) ~ pass_bechdel + log(budget_2013) + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
stepprofit <- stepAIC(modelprofit, direction = "both", trace = FALSE) %>% broom::tidy()
knitr::kable(stepprofit, digits = 3)
```
The algorithm selected budget, IMDB rating, and the genre variables of adventure, biography, drama, horror, music, romance, sci-fi, and western as influential predictors of a movie's profit. `pass_bechdel` was not selected as an influential predictor.

### IMDB Ratings

The second metric we assessed as a measure of movie success was movie reviews as measured by `imdb_rating`. Other potential predictors we included in the model were budget, `runtime`, `profit`, `award_winner`, and all genre variables.

```{r imdb, warning=FALSE}
# Stepwise regression for imdb ratings
## IMDB ~ Bechdel (binary) + logBudget + runtime + award_winner + profit + Genre
modelIMDB = lm(imdb_rating ~ pass_bechdel + log(budget_2013) + profit + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
stepIMDB <- stepAIC(modelIMDB, direction = "both", trace = FALSE) %>% broom::tidy()
knitr::kable(stepIMDB, digits = 3)
```
The algorithm selected the Bechdel test, budget, profit, and the genre variables of action, adventure, animation, biography, comedy, crime, drama, family, fantasy, horror, music, romance, thriller, and war as influential variables of a movie's IMDB rating. The p-value of `pass_bechdel` is significant at 5% and, on average, movies that pass the Bechdel test have a 0.255 point lower rating on IMDB than movies that do not pass.

### Budget

As budget is a significant predictor of both profit and IMDB ratings, we also assessed whether passing the Bechdel test is an influential predictor of a movie's budget. Other potential predictors we included in the model were `runtime` and all genre variables.

```{r budget log transform}
# Stepwise regression for budget
## logBudget ~ Bechdel (binary) + runtime + Genre
modelbudget = lm(log(budget_2013) ~ pass_bechdel + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
stepbudget <- stepAIC(modelbudget, direction = "both", trace = FALSE) %>%  broom::tidy()
knitr::kable(stepbudget, digits = 3)
```
The algorithm selected `runtime` and the genre variables of action, adventure, animation, biography, comedy, crime, documentary, drama, family, fantasy, musical, mystery, romance, sci-fi, sport, and thriller. `pass_bechdel` was not selected as an influential predictor.


In conclusion, based on our data, whether or not a movie passes the Bechdel test is not a significant predictor of that movie's potential for success based on the criteria of profit and budget. It is a significant predictor of a movie's IMDB ratings where movies that pass have, on average, a lower rating than movies that do not pass. Therefore, passing the Bechdel test can be considered detrimental to a movie's success.


## Scrap Code

```{r profit, include=FALSE}
# Stepwise regression for Profit
 
## Profit ~ Bechdel (binary) + budget + IMDB + Genre
model1 = lm(profit ~ pass_bechdel + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step1 <- stepAIC(model1, direction = "both", trace = TRUE)
summary(step1)
### bechdel selected, not significant

## Profit ~ Bechdel (binary) + budget + metascore + Genre
model2 = lm(profit ~ pass_bechdel + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step2 <- stepAIC(model2, direction = "both", trace = TRUE)
summary(step2)
### bechdel not selected 

## Profit ~ Bechdel (categorical) + budget + IMDB + Genre
model3 = lm(profit ~ bechdel_score + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step3 <- stepAIC(model3, direction = "both", trace = TRUE)
summary(step3)
### bechdel not selected

## Profit ~ Bechdel (categorical) + budget + metascore + Genre
model4 = lm(profit ~ bechdel_score + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step4 <- stepAIC(model4, direction = "both", trace = TRUE)
summary(step4)
### bechdel not selected

## logProfit ~ Bechdel (binary) + logBudget + metascore + Genre
model12 = lm(log(profit+1) ~ pass_bechdel + log(budget_2013) + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step12 <- stepAIC(model12, direction = "both", trace = TRUE)
summary(step12)
### bechdel not selected 

## logProfit ~ Bechdel (categorical) + logBudget + IMDB + Genre
model13 = lm(log(profit +1) ~ bechdel_score + log(budget_2013) + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step13 <- stepAIC(model13, direction = "both", trace = TRUE)
summary(step13)
### bechdel not selected

## logProfit ~ Bechdel (categorical) + logBudget + metascore + Genre
model14 = lm(log(profit +1) ~ bechdel_score + log(budget_2013) + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step14 <- stepAIC(model14, direction = "both", trace = TRUE)
summary(step14)
### bechdel not selected
```

```{r budget, include=FALSE}
## Budget ~ Bechdel (binary) + runtime + Genre
model9 = lm(budget_2013 ~ pass_bechdel + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step9 <- stepAIC(model9, direction = "both", trace = TRUE)
summary(step9)
### bechdel selected, not significant

## Budget ~ Bechdel (categorical) + runtime + Genre
model10 = lm(budget_2013 ~ bechdel_score + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step10 <- stepAIC(model10, direction = "both", trace = TRUE)
summary(step10)
### bechdel not selected

## Budget ~ Bechdel (categorical) + runtime + Genre
model101 = lm(log(budget_2013) ~ bechdel_score + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step101 <- stepAIC(model101, direction = "both", trace = TRUE)
summary(step101)
### bechdel not selected
```

```{r roi, include=FALSE}
# Stepwise Regression for Return on Investment (ROI)

## ROI ~ Bechdel (binary) + budget + IMDB + Genre
model5 = lm(ROI ~ pass_bechdel + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step5 <- stepAIC(model5, direction = "both", trace = TRUE)
summary(step5)
### bechdel not selected

## ROI ~ Bechdel (binary) + budget + metascore + Genre
model6 = lm(ROI ~ pass_bechdel + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step6 <- stepAIC(model6, direction = "both", trace = TRUE)
summary(step6)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + IMDB + Genre
model7 = lm(ROI ~ bechdel_score + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step7 <- stepAIC(model7, direction = "both", trace = TRUE)
summary(step7)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + metascore + Genre
model8 = lm(ROI ~ bechdel_score + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step8 <- stepAIC(model8, direction = "both", trace = TRUE)
summary(step8)
### bechdel not selected

```

```{r log transform roi, include=FALSE}
# Stepwise Regression for Return on Investment (ROI)

## logROI ~ Bechdel (binary) + budget + IMDB + Genre
model15 = lm(log(ROI+1) ~ pass_bechdel + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step15 <- stepAIC(model15, direction = "both", trace = TRUE)
step15 = step15 %>% broom::tidy()
knitr::kable(step15, digits = 3)
### bechdel not selected

## ROI ~ Bechdel (binary) + budget + metascore + Genre
model6 = lm(ROI ~ pass_bechdel + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step6 <- stepAIC(model6, direction = "both", trace = TRUE)
summary(step6)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + IMDB + Genre
model7 = lm(ROI ~ bechdel_score + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step7 <- stepAIC(model7, direction = "both", trace = TRUE)
summary(step7)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + metascore + Genre
model8 = lm(ROI ~ bechdel_score + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step8 <- stepAIC(model8, direction = "both", trace = TRUE)
summary(step8)
### bechdel not selected
```