---
title: "Stepwise Regression"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

```{r setup, include=FALSE}
library(tidyverse)
library(MASS)
knitr::opts_chunk$set(echo = TRUE)
# Read in dataset
movies_df <- read_csv("../Cathy/movies_df.csv")
```

##Using Stepwise Selection for Linear Regression: Is the Bechdel test a significant predictor of movie success?

In order to assess whether a movie passing the Bechdel test is a significant predictor of that movie's success, we utilized a stepwise selection algorithm. The `stepAIC` command from the `MASS` package performs stepwise model selection by optimizing AIC.

The setup of the regression models here follows the methodology outlined previously including log-transforming variables like `profit` and `budget_2013` to enforce a normal distribution.

### Profit

We first assessed the Bechdel test as a potential predictor of `profit`. Other potential predictors included in the model were `budget_2013`, `imdb_rating`, and all genre variables. Additionally, the binary Bechdel test variable was applied in these models to optimize degrees of freedom.

```{r profit log transfrom, warning=FALSE}
# Stepwise regression for Profit
## logProfit ~ Bechdel (binary) + logBudget + IMDB + Genre
model11 = lm(log(profit +1) ~ pass_bechdel + log(budget_2013) + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step11 <- stepAIC(model11, direction = "both", trace = FALSE) %>% broom::tidy()
knitr::kable(step11, digits = 3)
```
The algorithm selected budget, IMDB rating, and the genre variables of adventure, biography, drama, horror, music, romance, sci-fi, and western as influential predictors of a movie's profit. `pass_bechdel` was not selected as an influential predictor.

### Budget

As budget is a significant predictor of movie success (as assessed by profit - see table above), we also assessed whether passing the Bechdel test is an influential predictor of a movie's budget. Other potential predictors we included in the model are `runtime` and all genre variables.

```{r budget log transform}
## logBudget ~ Bechdel (binary) + runtime + Genre
model19 = lm(log(budget_2013) ~ pass_bechdel + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step19 <- stepAIC(model19, direction = "both", trace = FALSE) %>%  broom::tidy()
knitr::kable(step19, digits = 3)
```
The algorithm selected `runtime` and the genre variables of action, adventure, animation, biography, comedy, crime, documentary, drama, family, fantasy, musical, mystery, romance, sci-fi, sport, and thrill. `pass_bechdel` was not selected as an influential predictor.


In conclusion, based on our data, whether or not a movie passes the Bechdel test is not a significant predictor of that movie's potential for success based on the criteria of profit and budget.


## Scrap Code

```{r profit, include=FALSE}
# Stepwise regression for Profit
 
## Profit ~ Bechdel (binary) + budget + IMDB + Genre
model1 = lm(profit ~ pass_bechdel + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step1 <- stepAIC(model1, direction = "both", trace = TRUE)
summary(step1)
### bechdel selected, not significant

## Profit ~ Bechdel (binary) + budget + metascore + Genre
model2 = lm(profit ~ pass_bechdel + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step2 <- stepAIC(model2, direction = "both", trace = TRUE)
summary(step2)
### bechdel not selected 

## Profit ~ Bechdel (categorical) + budget + IMDB + Genre
model3 = lm(profit ~ bechdel_score + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step3 <- stepAIC(model3, direction = "both", trace = TRUE)
summary(step3)
### bechdel not selected

## Profit ~ Bechdel (categorical) + budget + metascore + Genre
model4 = lm(profit ~ bechdel_score + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step4 <- stepAIC(model4, direction = "both", trace = TRUE)
summary(step4)
### bechdel not selected

## logProfit ~ Bechdel (binary) + logBudget + metascore + Genre
model12 = lm(log(profit+1) ~ pass_bechdel + log(budget_2013) + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step12 <- stepAIC(model12, direction = "both", trace = TRUE)
summary(step12)
### bechdel not selected 

## logProfit ~ Bechdel (categorical) + logBudget + IMDB + Genre
model13 = lm(log(profit +1) ~ bechdel_score + log(budget_2013) + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step13 <- stepAIC(model13, direction = "both", trace = TRUE)
summary(step13)
### bechdel not selected

## logProfit ~ Bechdel (categorical) + logBudget + metascore + Genre
model14 = lm(log(profit +1) ~ bechdel_score + log(budget_2013) + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step14 <- stepAIC(model14, direction = "both", trace = TRUE)
summary(step14)
### bechdel not selected
```

```{r budget, include=FALSE}
## Budget ~ Bechdel (binary) + runtime + Genre
model9 = lm(budget_2013 ~ pass_bechdel + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step9 <- stepAIC(model9, direction = "both", trace = TRUE)
summary(step9)
### bechdel selected, not significant

## Budget ~ Bechdel (categorical) + runtime + Genre
model10 = lm(budget_2013 ~ bechdel_score + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step10 <- stepAIC(model10, direction = "both", trace = TRUE)
summary(step10)
### bechdel not selected

## Budget ~ Bechdel (categorical) + runtime + Genre
model101 = lm(log(budget_2013) ~ bechdel_score + runtime + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step101 <- stepAIC(model101, direction = "both", trace = TRUE)
summary(step101)
### bechdel not selected
```

```{r roi, include=FALSE}
# Stepwise Regression for Return on Investment (ROI)

## ROI ~ Bechdel (binary) + budget + IMDB + Genre
model5 = lm(ROI ~ pass_bechdel + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step5 <- stepAIC(model5, direction = "both", trace = TRUE)
summary(step5)
### bechdel not selected

## ROI ~ Bechdel (binary) + budget + metascore + Genre
model6 = lm(ROI ~ pass_bechdel + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step6 <- stepAIC(model6, direction = "both", trace = TRUE)
summary(step6)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + IMDB + Genre
model7 = lm(ROI ~ bechdel_score + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step7 <- stepAIC(model7, direction = "both", trace = TRUE)
summary(step7)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + metascore + Genre
model8 = lm(ROI ~ bechdel_score + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step8 <- stepAIC(model8, direction = "both", trace = TRUE)
summary(step8)
### bechdel not selected

```

```{r log transform roi, include=FALSE}
# Stepwise Regression for Return on Investment (ROI)

## logROI ~ Bechdel (binary) + budget + IMDB + Genre
model15 = lm(log(ROI+1) ~ pass_bechdel + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step15 <- stepAIC(model15, direction = "both", trace = TRUE)
step15 = step15 %>% broom::tidy()
knitr::kable(step15, digits = 3)
### bechdel not selected

## ROI ~ Bechdel (binary) + budget + metascore + Genre
model6 = lm(ROI ~ pass_bechdel + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step6 <- stepAIC(model6, direction = "both", trace = TRUE)
summary(step6)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + IMDB + Genre
model7 = lm(ROI ~ bechdel_score + budget_2013 + imdb_rating + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step7 <- stepAIC(model7, direction = "both", trace = TRUE)
summary(step7)
### bechdel not selected

## ROI ~ Bechdel (categorical) + budget + metascore + Genre
model8 = lm(ROI ~ bechdel_score + budget_2013 + metascore + action + adventure + animation + biography + comedy + crime + documentary + drama + family + fantasy + history + horror + music + musical + mystery + romance + sci_fi + sport + thriller + war + western, data = movies_df)
step8 <- stepAIC(model8, direction = "both", trace = TRUE)
summary(step8)
### bechdel not selected
```